Fast Sorting on Modern Computers

### 2005-08-27 docs/tech_report/conclusions.tex
started making fixes to conclusions

We found that algorithm S has a lower instruction count than algorithm N, but
that this does not necessarily translate to better performance. Our results show
also that the level 1 cache needs to be taken into account, that the extra
instruction in multi-mergesort make it slower than double-aligned tiled
mergesort, and that multi-mergesort has regular branch predictive access
patterns during the k-way merge, which can be taken advantage of by a two-level
adaptive predictor.

@@

Our results show base quicksort being the fastest of the quicksorts. We note our
lowest instruction count occurs with the median-of-7 partitioning. We also see
that using a binary search slows multi-quicksort down compared to a sequential
search, but that its extra instructions make it slower than memory-tuned
quicksort.  Despite being slower in our tests, our simulations show tuned
quicksort to have fewer cache misses, instructions and branch mispredictions
than base quicksort.

@@

\section{Branch Results}
%TODO this section has to be reread by David

@@

Insertion sort has exactly one miss per key. Using this result we were able to
make several other sorts faster by replacing their simple inline sorts with
insertion sorts.

Selection sort also has a very low number of branch mispredictions, the results
of which match up well to Knuth's calculation. We found as well that bubblesort
and shakersort have incredibly bad branch prediction performance. Both are
$O(N^2)$ sorts, and the number of branches and branch misses have the same
complexity.
%TODO i dont like how this reads

@@

In heapsort, we found a pattern of branch mispredictions when using a fanout in
our 8-heap, and that this pattern matched closely that of selection sort,
discussed above. We found that the approximated results using $H_N$ was similar to
our expected results for large arrays, and followed a similar pattern for small
arrays.

@@

\section{Important Results and Contributions}

%TODO lamarca does an overall comparison. we should do the same

\section{Instruction Count Results}

>>>>

We found that algorithm S has a lower instruction count than algorithm N, but
that this does not necessarily translate to better performance. Our results show
also that the level 1 cache needs to be taken into account, that the extra
instruction in multi-mergesort make it slower than double-aligned tiled
mergesort, and that multi-mergesort has regular access patterns during the k-way
merge, which can be taken advantage of by a two-level adaptive predictor.

@@

Our results show base quicksort being the fastest of the quicksorts. We note our
lowest instruction count occurs with the median-of-three partitioning. We also
see that using a binary search slows multi-quicksort down compared to a
sequential search due to branch mispredictions, but that its extra instructions
make it slower than memory-tuned quicksort. Despite being slower in our tests,
our simulations show memory-tuned quicksort to have fewer cache misses,
instructions and branch mispredictions than base quicksort.

@@

\section{Branch Results}

@@

\subsection{Elementary Sorts}
Insertion sort has exactly one miss per key. Using this result we were able to
make several other sorts faster by replacing their simple inline sorts with
insertion sorts.

Selection sort also has a very low number of branch mispredictions, the results
of which match up well to Knuth's calculation. We found as well that bubblesort
and shakersort have incredibly bad branch prediction performance. Both are
$O(N^2)$ sorts, and the number of branches and branch misses have the same
complexity.

@@

\subsection{Heapsort}
In heapsort, we found a pattern of branch mispredictions when using a fanout in
our 8-heap, and that this pattern matched closely that of selection sort,
discussed above. We found that the approximated results using $H_N$ was similar to
our expected results for large arrays, and followed a similar pattern for small
arrays.

@@
